# Reproducible Research: Programming Assignment 1

## Loading and preprocessing data
```{r loaddata}
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
total_Steps <- tapply(data$steps, data$date, sum, na.rm = FALSE)
qplot(total_Steps,bindwidth = 1000, xlab = "total number of steps taken per day")
mean(total_Steps, na.rm = TRUE)
median(total_Steps, na.rm = TRUE)
```

## What is the average daily activity pattern?
```{r}
library(ggplot2)
averages <- aggregate(x = list(steps = data$steps), by= list(interval= data$interval), FUN =mean, na.rm = TRUE)
p <- ggplot(data = averages, aes(x=interval, y=steps)) 
p + geom_line() + xlab("5-min interval") + ylab ("average number of steps taken")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
averages[which.max(averages$steps),]
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
missing <- is.na(data$steps)
table(missing)
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval,etc

```{r}
# Replace missing value with mean value for that 5-min interval
fill <- function(interval, steps){
  filled <- NA
  if (!is.na(steps))
    filled <- c(steps)
  else 
    filled <- averages[averages$interval == interval, "steps"]
  return(filled)
}
newdata <- data
newdata$steps <- mapply(fill, newdata$interval, newdata$steps)
```

Using the new data set, let's make a histogram of the total number of steps taken each data and calculate the mean and median total number of steps

```{r}
total_Steps <- tapply(newdata$steps, newdata$interval, sum)
qplot(total_Steps, binwidth = 1000, xlab ="total number of steps taken each day")
mean(total_Steps)
median(total_Steps)
```

#Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday_weekend <- function(date) {
  day <- weekdays(date)
  if (day %in% c("Monday","Tuesday", "Wednesday", "Thursday","Friday"))
    return("weekday")
  else if (day %in% c("Saturday","Sunday"))
    return("weekend")
  else
    stop("Invalid date")
}
newdata$date <- as.Date(newdata$date)
newdata$day <- sapply(newdata$date, weekday_weekend)
```

```{r}
# Make a panel plot containing plots of average number of steps taken on weekdays and weekends.
averages <- aggregate(steps ~ interval + day, data = newdata, mean)
p <- ggplot(averages, aes(interval, steps))
p + geom_line() + facet_grid(day ~.) + xlab("5-min interval") + ylab("Number of steps")
```